Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors

### 2012-12-17 Publications/2013-IPMI/Latex/045-Experiments.tex
reordering / adding experiments
work in progress

\section{Data and experiments}
\label{sec:experiments}
%
\subsection{Simulated datasets}
%
As described in \autoref{sec:introduction}, the general situation in
the connectivity pipelines consists on having 
a reliable segmentation obtained from the high resolution \ac{t1} 
reference image. Therefore, a precise location of the tissue interfaces
of interest is available in a reference space. Given that there is no 
interest on the anatomical reference segmentation,
we directly obtained the shape priors from the models. \\

On the other hand, the target \ac{dwi} data is characterized by its
low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
posterior reconstruction methodology and the angular resolution
intended, the \ac{dwi} raw data has to be processed in order to
extract the information in a manageable manner. Particularly, we
will use the \ac{fa} and \ac{md} maps for convenience.
Whereas \ac{fa} describes the \emph{shape} of diffusion, 
the \ac{md} depicts the \emph{manitude} of the process. 
There exist two main reasons to justify their choice. 
First, they are well-understood and standardized in clinical routine.
Second, together they contain most of the information that is
usually extracted from the \ac{dwi}-derived scalar maps. \\

In order that demonstrating the functionality of the proposed
methodology and characterize its possibilities, we developed two
synthetic models and simulated their corresponding \ac{dwi}
raw signal as described in \citep{tuch_q-ball_2004}. 
(HERE WE NEED A GOOD DESCRIPTION OF THE DATA, directions, resolution, etc)
. The first model consists of several spherical shapes emulating
the different brain tissues (see \autoref{fig:fa}, first row). 
The second model is based on the BrainWeb dataset.
We reconstructed the \ac{dwi} signal with standard procedures to approximate the
environment to the real one at maximum. \\

%
\subsection{Experiment}
%
For both models, we created manually a sound distortion visually similar
to real \ac{epi} distortions. We interpolated the distortion to a 
dense deformation field, necessary for warping the raw \ac{dwi} simulated
data. Once the signal was deformed, we proceeded to reconstruct the
\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.\\

We evaluate the performance of our methodology to estimate the deformation
field and compare it to the original synthetic deformation field applied to the data.

>>>>

\section{Data and experiments}
\label{sec:experiments}
%
\subsection{Shape prior}
%
As described in \autoref{sec:introduction}, the general situation in
the connectivity pipelines consists on having 
a reliable segmentation obtained from the high resolution \ac{t1} 
reference image. Therefore, a precise location of the tissue interfaces
of interest is available in a reference space. Given that the anatomical reference segmentation is beyond the scope of this manuscript, we simply rely on the shape priors obtained from the models. \\

%
\subsection{Synthetic gray-scale data}
%
The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}. \textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.

%
\subsection{Simulated diffusion data}
%
In order to demonstrate the specific functionality of the proposed
methodology and characterize its possibilities with diffusion data, we developed a
synthetic model and simulated its corresponding \ac{dwi}
raw signal as described in \citep{tuch_q-ball_2004}.

This second model consists of several spherical shapes emulating
the different brain tissues (see \autoref{fig:fa}, first row). 
We reconstructed the \ac{dwi} signal with standard procedures to approximate the
environment to the real one at maximum. \\

The target \ac{dwi} data is characterized by its
low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
posterior reconstruction methodology and the angular resolution
intended, the \ac{dwi} raw data has to be processed in order to
extract the information in a manageable manner. Here, we
will use the \ac{fa} and \ac{md} maps for convenience.
Whereas \ac{fa} describes the \emph{shape} of diffusion, 
the \ac{md} depicts the \emph{magnitude} of the process. 
There exist two main reasons to justify their choice. 
First, they are well-understood and standardized in clinical routine.
Second, together they contain most of the information that is
usually extracted from the \ac{dwi}-derived scalar maps. \\

For this model, we created manually a sound distortion visually similar
to real \ac{epi} distortions. We interpolated the distortion to a 
dense deformation field, necessary for warping the raw \ac{dwi} simulated
data. Once the signal was deformed, we proceeded to reconstruct the
\ac{dti} \textbf{FIXME is it really DTI, here, not rather DSI?} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.

\textbf{FIXME Give parameters here?}\\
